---
title: "Panel Survey Data Workshop"
author: "Norah Rami"
editor: visual
---

```{r}
library(ggplot2)
library(readr)
library(ggdag)
library(tidyverse)
library(gt)
library(modelsummary)
library(fixest)

# read-in data
dat = read_csv("https://raw.githubusercontent.com/jrspringman/psci3200-globaldev/main/workshops/aau_survey/clean_endline_did.csv" ) %>%
    # clean home region variable
  mutate(q8_baseline = ifelse(q8_baseline == "Southern Nations, Nationalities, and Peoples Region", "SNNPR", q8_baseline), 
         q8_baseline = str_remove(q8_baseline, " Region"))
# create color palette for plotting
palette = MetBrewer::met.brewer(name = "Cross")
```

```{r}
#In this function, I select for the baseline and endline of all the variables we are interested in, renaming them to more descriptive variable names.
panel_survey <- dat |> select(response_id, treatment_status, user_language, gender = q3_baseline, civsoc_career_bl = q26_civ_baseline, civsoc_career_el = q26_civ,  politics_career_bl = q26_politics_baseline, politics_career_el = q26_politics, pubsec_career_bl = q26_public_baseline, pubsec_career_el = q26_public, office_plan_bl = q27_1_baseline, office_plan_el = q27_1, ngo_plan_bl = q27_3_baseline, ngo_plan_el = q27_3, belief_self_poschange_bl = q17_3_baseline, belief_self_poschange_el = q17_3, belief_youthengage_bl = q17_1_baseline, belief_youthengage_el = q17_1, belief_youth_poschange_bl = q17_2_baseline, belief_youth_poschange_el = q17_2)
```

# Creating Index Variables

## Additive v. Index Variables

An additive index is the cumulative value of other variables by simply adding them together. Thus, for an additive index to be significant, it's important that the variables that are being combined are on the same scale. Additive index variables are useful when we're interested in the cumulative amount of something.

An average Z-scores refers to an average of a set of z-scores across variables. A z-score is a standardized measurement based on a mean of 0 and a standard deviation of 1. Averaged z-scores are useful when variables are measured on different scales or when variables cannot be summed.

```{r}
panel_survey <- panel_survey %>%
#creating the additive index variable
  mutate(
    index_civsoc_pubsec_plan_bl = as.numeric(pubsec_career_bl) + 
                                as.numeric(civsoc_career_bl), index_civsoc_pubsec_plan_el = as.numeric(pubsec_career_el) + as.numeric(civsoc_career_el)
  ) %>%
#creating the z-score versions of each variable 
  mutate(
    z_civsoc_pubsec_plan_bl = scale(as.numeric(index_civsoc_pubsec_plan_bl)), 
    z_civsoc_pubsec_plan_el = scale(as.numeric(index_civsoc_pubsec_plan_el)), 
    z_pol_efficacy_bl = scale(rowMeans(cbind(
      belief_youth_poschange_bl,
      belief_self_poschange_bl,
      belief_youthengage_bl
    ), na.rm = TRUE)), 
    z_pol_efficacy_el = scale(rowMeans(cbind(
      belief_youth_poschange_el,
      belief_self_poschange_el,
      belief_youthengage_el
    ), na.rm = TRUE))
                                          
  )
```

```{r}
#visualizing the code
# Public Sector Career (Baseline)
ggplot(panel_survey, aes(x = z_civsoc_pubsec_plan_bl)) +
  geom_density(fill = "#3182bd", alpha = 0.6) +
  labs(
    title = "Distribution of Z-Scores: Future plans for a career in public sector or civil society (Baseline)",
    x = "Z-score",
    y = "Density"
  ) +
  theme_minimal()

# Civil Society Career (Baseline)
ggplot(panel_survey, aes(x = z_civsoc_pubsec_plan_el)) +
  geom_density(fill = "#31a354", alpha = 0.6) +
  labs(
    title = "Distribution of Z-Scores: Future plans for a career in public sector or civil society (Endline)",
    x = "Z-score",
    y = "Density"
  ) +
  theme_minimal()

# Public Sector Career (Endline)
ggplot(panel_survey, aes(x = z_pol_efficacy_bl)) +
  geom_density(fill = "#9ecae1", alpha = 0.6) +
  labs(
    title = "Distribution of Z-Scores: Feelings of political efficacy (Baseline)",
    x = "Z-score",
    y = "Density"
  ) +
  theme_minimal()

# Civil Society Career (Endline)
ggplot(panel_survey, aes(x = z_pol_efficacy_el)) +
  geom_density(fill = "#a1d99b", alpha = 0.6) +
  labs(
    title = "Distribution of Z-Scores: Feelings of political efficacy (Endline)",
    x = "Z-score",
    y = "Density"
  ) +
  theme_minimal()
```

According to these graphs, the treatments didn't result in significant changes in feelings of political efficacy or future career plans. However, we can see a slight increase in respondents post-treatment who wanted to pursue a career in the public sector or civil society as well as an increase in beliefs in public efficacy.

# Estimating Models

```{r}
panel_survey <- panel_survey %>%
  select(response_id, gender, z_civsoc_pubsec_plan_bl, z_civsoc_pubsec_plan_el, z_pol_efficacy_bl, z_pol_efficacy_el)
model <- lm(z_civsoc_pubsec_plan_bl ~ z_pol_efficacy_bl, data = panel_survey)

modelsummary(model,
             stars = TRUE,
             title = "Regression of Future Plans on Political Efficacy (Baseline)",
             output = "markdown")
```

## Interpretation

$\alpha$ When political efficacy is at zero (which is the sample mean of the political efficacy index) the predicted future plans for civil society/public sector is 0.001

$\beta_1$This means that higher political efficacy at baseline is positively and significantly associated with stronger future plans for civil society/public sector work.

This model asserts that beliefs in political efficacy is related to higher aspirations to work in civil society and public sector. This assumes that there is not a confounding variable that influences both political efficacy and civil society/public sector. In realty however, education, class, and gender can all impact both one's feelings of political efficacy as well as their career prospects.

# Regression 2:

```{r}

#creating a binary variable using if else 
panel_survey <- panel_survey %>%
  mutate(
    b_pol_efficacy_bl = ifelse(z_pol_efficacy_bl >= mean(z_pol_efficacy_bl, na.rm = TRUE), 1, 0),
    b_pol_efficacy_el = ifelse(z_pol_efficacy_el >= mean(z_pol_efficacy_el, na.rm = TRUE), 1, 0)
  )

#linear regression with interaction terms

model_interaction <- lm(z_civsoc_pubsec_plan_bl ~ b_pol_efficacy_bl * gender, data = panel_survey)

# Display with modelsummary
library(modelsummary)

modelsummary(model_interaction,
             stars = TRUE,
             title = "Interaction Between Political Efficacy and Gender (Baseline)",
             output = "markdown")
```

## Interpretation

$\alpha$: The expected future plans score fore women with low political efficact is slightly below the mean at the baseline

$β1 = 0.224$ is the effect of political efficacy on future plans at the baseline This indicates that for women (which we are considering the base gender) a 1-unit increase in political efficacy is associated with a positive and marginally significant increase (at the 10% level) which suggests political efficacy has a positive relationship with future plans, but its not strongly conclusive

$\beta2 = 0.008$ This coefficient reflects the effect of gender on future civil society/public sector plans when poliical efficacy is held constant. The coefficient is not significant, meaning that gendere does not have a significant effect on future plans in this model.

$\beta_3 = 0.025$ the interaction term shows how the relationship between political efficacy and future plans differs for men in comparison to women. The coefficient of **0.025** suggests that **men** with higher political efficacy are **slightly more likely** to have increased future plans compared to females, but the effect is **not statistically significant**.

However, the model fit is relatively weak, explaining only a small fraction of the variance. Gender it seems does not inform the relationship between political efficacy and career goals.

# Long Panel

```{r}

#creating a long panel that creates two rows for each response_id. all variables that end with _bl (indicating baseline) and _el (indicating endline) are split into a second row. a new variable called wave indicates whether the data is from the endline wave or the baseline wave
panel_long <- panel_survey %>%
  pivot_longer(
    cols = contains("z_") | contains("b_"),
    names_to = c(".value", "wave"),
    names_pattern = "(.*)_(bl|el)"
  ) %>%
  mutate(wave = ifelse(wave == "bl", "baseline", "endline"))

#creating a fixed effects model

mod_fe <- feols(
  z_civsoc_pubsec_plan ~ b_pol_efficacy | response_id,  # pipe syntax for fixed effects
  data = panel_long,
  cluster = ~response_id
)

modelsummary(mod_fe, stars = TRUE, title = "Fixed Effects: Political Efficacy and Career",
             output = "markdown")

```

## Interpretation:

β1 = -0.058

This tells us that, **within individuals over time**, those whose political efficacy increased above the sample mean had a **0.058 standard deviation *decrease*** in their future plans index.

The sign is **negative**, which is counterintuitive if we expected political efficacy to increase interest in public/civic careers, based on our past models

However, the **standard error** is **0.057**, and the coefficient is **not statistically significant** (no stars), meaning this effect could be due to chance.

By applying fixed effects, when we control for the individual we don't see a positive nor significant correlation between political efficacy and career aspirations in civil society or the public sector.
